---
title: "Explore size spectra"
author: "Asta, Freddie, Shane, Ken"
date: "07/02/2021"
output: html_document
---

### libraries

```{r setup, include=FALSE}
#rm(list=ls())
library(tidyverse)
library(data.table)
```

### Theoretical size spectrum 

All code and data for the figures in: K.H. Andersen (2019) "Fish Ecology, Evolution, and Exploitation", https://press.princeton.edu/titles/13516.html are on GitHub: https://github.com/Kenhasteandersen/Fish

To reproduce population size spectrum (Fig.4.2) we need plotSpectrum() in plotChapterDemography.R

To run the function we need the package:  https://github.com/Kenhasteandersen/FishSizeSpectrum

```{r, eval = F}
#1: either install the package from github
library(devtools)
install_github("Kenhasteandersen/FishSizeSpectrum", ref = "master")
library(fishsizespectrum)
## NOTE - if for some reason you cannot install this package and run the rest of this chunk, the datasets it produces are saved at the end of it
```

```{r}
p <- baseparameters()   ## setup basic parameters
W <- 1000  ## maximum body size (Winf)
p$W <- W
N <- spectrum(p)  #get the spectrum based on biphasic growth numerical simulation
Nana <- spectrumana(p) #get the spectrum for analytical solution using Von-Bertanfy growth curve

# write these spectra into csv files
N %>% as_tibble() %>% select(-c(mu.1:mu.1000)) %>% fwrite("theoryBase_biphasic.csv")
Nana %>% as_tibble() %>% fwrite("theoryBase_anal.csv")

```

This spectrum is from Andersen et al. 2019, Figure 4.1a. Details on how the assumptions are explained in the reference. Here we just get the data to plot the figure

```{r cars}
model1 <- read_csv("theoryBase_biphasic.csv")  # assuming biphasic growth
model2 <- read_csv("theoryBase_anal.csv")  # assuming VB growth 

theory <- ggplot() + 
  geom_line(data = model1, aes(x = w, y = NprR/NprR[1]), size=1) +
  geom_line(data = model2, aes(x = w, y = NprR/NprR[1]), size=1, col = 'red') +
  scale_y_log10() +
  scale_x_log10() + 
  xlab(expression(log10(w))) +
  ylab(expression(log10(Ndens/Ndens[smallest]))) +
  theme_bw() + 
  ggtitle("Theoretical number DENSITY spectrum")

theory
```

## Empirical size spectra

We will use data from RLS for Jervis Bay marine protected area, all surveys combined

```{r}
#read in data
df_single <- read_csv("jervisBay.csv")

cutoff = c(2.5, 5, 7.5, 10, 12.5, 15, 20, 25, 30, 35, 40, 50, 
    seq(from=62.5, to=200, by=12.5), 
    seq(from=250, to=500, by=50))

rls_bins = c(0, cutoff)

df_cutoff <- tibble(
  size_class = cutoff,
    size_indx = 1:length(cutoff)
)

df_single <- left_join(df_single, df_cutoff, by = "size_class")

df_abundance <- df_single %>%
    group_by(spp_id) %>%
    summarise(.groups = "drop", Total_N = n(), Mean_L = mean(size_class))

df_fit <- df_single %>%
    group_by(spp_id, size_indx, w) %>%
    summarise(.groups = "drop", n = n()) %>%
    left_join(df_cutoff, by = "size_indx") %>%
    left_join(df_abundance, by = "spp_id")

## plot raw length / abundance observations for species that have more than 20 observations
ggplot(data = df_fit %>% filter(Total_N > 20)) +
    geom_point(mapping = aes(x = size_class, y = n)) + 
    geom_line(mapping = aes(x = size_class, y = n)) +
    facet_wrap( ~ spp_id, ncol = 5) +
    scale_y_log10() + scale_x_log10() +
    labs(x = "Fish length (cm)",y = "Fish abundance") +
    theme_bw()


## Now do the same plot but scaled by mean length 
df_plot <- df_fit %>% 
    mutate(l_hat = log(size_class) - log(Mean_L))

ggplot(df_plot, aes(x = l_hat, y = n/Total_N, color = log(Total_N), group = spp_id)) + 
    geom_line() +
    scale_y_log10() +
     scale_colour_gradient2(low = "#fed976", mid = "#fd8d3c", high = "#bd0026", 
         midpoint = 2.5, na.value = NA) +
    theme_bw()

save(df_plot, file = "emp_data_plot.RData")

```

Within species variation in the log-size distributions are similar in shape (but not location) between species, and negative quadratic, supporting our assumption of a normal distribution: li,s∼N(l~s,ϕwl~s).

### rescale theoretical data 

To compare theoretical data with empirical observations we need to put theoretical predictions into same size bins, as in empirical data 

#### model 1

```{r, eval = F}
### Step 1
### convert theoretical data to Numbers per size bin and not number density per size bin
### for this we need to multiply number density by the width of the size bin 


#get width of the size bins dw, which is the difference between two consecutive size classes. At the moment they are not logarithmic 
model1$dw <- NA #bin width

for (i in 2:length(model1$w)) {
  model1$dw[i-1] <- model1$w[i] - model1$w[i-1]
}

## now to get actual numbers we multiply number density (NprR) by dw
model1$numb <- model1$NprR * model1$dw
model1 <- model1[c(1:999),]
sum(model1$numb)

### Step 2
## now we convert theoretical weights into lengths using standard LW conversion data
LWa = 0.01
LWb = 3
model1$len <- (model1$w/LWa)^(1/LWb)

## Now to convert number density into numbers per bin I will use an clumsy method which still works 
### Step 3
# convert numbers per small unit volume (m3?) into larger number (per 10km2) by say multiplying numbers by 10e7 and rounding to 0. This will give numbers as digits and not real values
model1$numb2 <- round(model1$numb*1e7, 0) ## 

sum(model1$numb2) # looks like a large enough number, so should be a sufficient approximation

# untable the data so each row represents one theoretical observation of a certain length
library(reshape)
temp_size <- untable(model1, num = model1$numb2)

### Step 4
# bin these obsevations into same lengths groups as RLS data
m1_plot <- hist(temp_size$len, breaks = c(0,cutoff), plot = F)

teor_plot <- as.data.frame(cbind(m1_plot$mids, m1_plot$counts))
colnames(teor_plot) <- c("size_class", "n")

#get mean length (based on abundance, like in RLS), max L and total abundance
meanL <- mean(temp_size$len)
maxL <- max(temp_size$len)
totalN <- length(temp_size$len)

#rescale by mean length 
teor_plot$l_hat <- log(teor_plot$size_class) - log(meanL)
#get relative numbers (divided to sum)
teor_plot$rel_N <- teor_plot$n/totalN

teor_plot1 <- teor_plot
save(teor_plot1, file = "teor_plot_model1c.RData")

```

#### model2
do exactly the same with model 2 data

```{r, eval = F}
#rename the file, so I can use the same script 
model1 <- model2

### Step 1
### convert theoretical data to Numbers per size bin and not number density per size bin
### for this we need to multiply number density by the width of the size bin 

#get width of the size bins dw, which is the difference between two consecutive size classes. At the moment they are not logarithmic 
model1$dw <- NA

for (i in 2:length(model1$w)) {
  model1$dw[i-1] <- model1$w[i] - model1$w[i-1]
}

## now to get actual numbers we multiply number density by dw
model1$numb <- model1$NprR * model1$dw
model1 <- model1[c(1:999),]
sum(model1$numb)

### Step 2
## now we convert theoretical weights into lengths using standard LW conversion data
LWa = 0.01
LWb = 3

model1$len <- (model1$w/LWa)^(1/LWb)

### Step 3
model1$numb2 <- round(model1$numb*1e7, 0) ## 

sum(model1$numb2)

# untable the data so each row represents one theoretical observation of a certain length
library(reshape)
temp_size <- untable(model1, num = model1$numb2)

### Step 4
# bin them into same lengths as RLS data
m1_plot <- hist(temp_size$len, breaks = c(0,cutoff),plot = F)

teor_plot <- as.data.frame(cbind(m1_plot$mids, m1_plot$counts))
colnames(teor_plot) <- c("size_class", "n")

#get mean, max L and total abundance
meanL <- mean(temp_size$len)
maxL <- max(temp_size$len)
totalN <- length(temp_size$len)

#rescale by mean length 
teor_plot$l_hat <- log(teor_plot$size_class) - log(meanL)
teor_plot$rel_N <- teor_plot$n/totalN

teor_plot2 <- teor_plot
save(teor_plot2, file = "teor_plot_model2c.RData")

```

### Fig1 plot

```{r}
load(file = "teor_plot_model1c.Rdata") #biphasic growh
load(file = "teor_plot_model2c.Rdata") #VB analytical solution
load(file = "emp_data_plot.RData")

#first check empirical data plot
fig_number <- 
  df_plot %>% 
  ggplot(aes(x = l_hat, y = n/Total_N)) + 
  geom_line(aes(color = log(Total_N), group = spp_id)) +
  xlim(-2, 2) +
  scale_y_log10() +
  #scale_x_log10() +
  scale_colour_gradient2(low = "#fed976", mid = "#fd8d3c", high = "#bd0026", 
                         midpoint = 2.5, na.value = NA) +
  theme_bw()

fig_number

#now check theoretical data plot
teor_plot1 <- teor_plot1[c(which(teor_plot1$n > 0)),]
teor_plot2 <- teor_plot2[c(which(teor_plot2$n > 0)),]

fig_teor <- ggplot() +
  geom_line(data=teor_plot1, aes(x=l_hat, y=rel_N), size=1) +
  geom_line(data=teor_plot2, aes(x=l_hat, y=rel_N), size=1, col = 'red') + 
  xlim(-2, 2) +
  scale_y_log10() +
  theme_bw()
fig_teor

# Combined fig
p1 <- 
  fig_number +
  geom_smooth(data=teor_plot1, aes(x=l_hat, y=rel_N), se = F, size=1.2, col = "navy") + ## biphasic
  geom_smooth(data=teor_plot2, aes(x=l_hat, y=rel_N), se = F, size=1.2, col = "blue") +  #VB analytical
  xlab("log10(length/mean_length)") +
  ylab("Relative numbers") +
  theme_bw(24) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        #legend.title = element_blank(),
        legend.position = "none")

p1

```


